\chapter{Results and Discussion}
\label{ch:Discussion}
Now that it is understood how the study was conducted and what the architecture of the final prototype is, the results of the conducted Action Research and the state of the prototype can be categorically discussed. The results mainly reference the final review from chapter \ref{ch:finalReview}, but will also take a holistic look at the developed prototype and compare it to its contemporaries.

\section{Knowledge Generation and Transparency Mechanisms}
\label{sub:knowledgeGenerationMechanism}
The first category deals with the way that \prototype{} generates its answers and how to achieve the best results when interacting with the system.

Differences in prompting style greatly influenced output characteristics. Ambiguous prompts, without domain-specific language and terms, left interpretation up to the LLM and often returned results which did not align with the expectations of the experts. However, when prompts were fine tuned to resemble expert terminology and domain language, retrieval more frequently referenced elements in the knowledge graph containing the expected information. This is because the LLM has more context to build the Cypher with. For example, when supplementing the prompt with specific key words, the LLM is able to generate a Cypher containing these key words. The observed results exemplify how crucial the quality of prompting is in order to achieve a desired result.

Applied to real-world scenarios, this highlights two types of users who can benefit from this prompting strategy. First, it is an advantage if the user is aware of the details of the underlying knowledge graph, as this enables the user to prompt specifically for the knowledge graph at hand. If a user is aware that the knowledge graph was built off imported Archi models and understands the schema, then they will be able to add context to the prompt to retrieve the information that they require to answer their question. A second user group	 who may achieve better results with the proprietary knowledge graph of the final review is users who understand the technical jargon underlying the implementation, in this example the employees of the examination office. Even without the necessary technical knowhow to prompt the system for keywords found in the knowledge graph, they are able to apply their own specialized terminology within the prompt. This specific context may help the LLM to generate the correct Cyphers to assist in such a use case. In contrast, a user who is neither aware of the underlying implementation nor knows the technical jargon of the examination office will struggle much more to achieve the desired result when interacting with the system.

This ties back to section \ref{sec:intro:relatedWork}, where Zhao et al. (2024) \cite{zhao2024retrieval} describe that the context given to the LLM allows it to uncover hidden rationales, allowing it to learn through the provided context. This implies that the user must also be active during the interaction with the system and may not fall into a naïve usage pattern if they are to extract the most value out of the system.

Next, the common challenge of LLMs hallucinating when generating answers could not be mitigated during this research. In the example of section \ref{sec:observedSysBehavior} the observed reference to the QIS system illustrates this problem. Although QIS is not mentioned in the department's architecture, it is output and described by the LLM as if it were, with no indication of uncertainty. This is a clear case in which the general knowledge on which an LLM is pretrained is leaked into the generated answer. Hallucinations such as this may go unnoticed and although minuscule in the observed case, the effects this may have in a real-world scenario could be large. The result of this is that the user must be aware that the LLM may hallucinate its answers, requiring them to constantly fact check and ensure the correctness of outputs. When dealing with EAM data via such a RAG-based system, it is a requirement to be able to separate graph-grounded knowledge from the model's trained knowledge. % TODO: Elaborate on this more. this is an important point. could this be automated? could a solution be that only keywords found in the database is used?

This leads to the key advantage that \prototype{} has over commercial LLM tools such as Rovo which is its transparency in how answers are generated. \prototype{} allows the user to copy and manually execute the exact Cypher query used to retrieve the data in Neo4j, making the underlying information directly inspectable. In contrast, Rovo does not provide insight into how its responses are derived. This inspectability enables users to validate outputs against the knowledge graph and detect potential hallucinations. The distinction therefore lies between a black-box system such as Rovo and a grey-box system such as \prototype{}, where the reasoning path remains transparent.

In a real-world setting, this added transparency could make a difference when discussing how to implement an LLM in regard to data integrity and risk management. The enabled transparency allows for the data do be double-checked, ensuring that the data's integrity is upheld. At the same time, the risk of introducing misinformation into an enterprise's architecture may be reduced this way.

The challenge of dealing with hallucinations goes beyond the scope of \prototype{}, as Rovo also produced factually incorrect answers. This comparison between \prototype{} and Rovo brings to light the necessity for features that enable such a level of transparency. This is also confirmed in the paper by Zhao et al. (2024), where the authors mention that "enhancing transparency in LLMs and reducing the uncertainty of the generated output is critical for increasing trust and reliability to their outputs, especially in fields where precision and accountability are paramount"\cite{zhao2024retrieval} such as in the field of EAM.

The above points were a common theme throughout the development and final review. They show that having an understanding of the system at hand can give the user an advantage to extract the most value from it as well as to mitigate errors.

\section{Architectural Limitations}
\label{sub:architecturalLimitations}
The results of the final review identified several technical limitations in \prototype{}'s current version.

A learning moment presented itself during the final review when \prototype{} was revealed to have tricked the Cypher generation process. The Cypher queries containing a templated answer showcase how relevant the underlying system prompt is. The prompt enables the LLM to complete the task, but does not ensure correctness of the completed task. The predefined system prompt tasks the LLM with generating one Cypher per user request. It is supplemented with information on how to complete the task as well as some pointers on how to structure the Cypher. However, the system prompt never explicitly tells the LLM to not hardcode any self-generated information into the Cypher. This bypasses the knowledge extraction from the knowledge graph and essentially executes an empty Cypher. The result is that the LLM satisfies the given instruction by generating a Cypher and using the results of the Cypher to generate the final response. But it does not do so in a meaningful way by extracting relevant information from the knowledge graph. This highlights that the LLM may be following instructions, but it is not following the intent of the system prompt. The LLM itself may be optimized for instruction satisfaction but not task correctness.

Although this loophole within the system prompt was abused by the LLM, catching this behavior is an attest to the transparency advantage of \prototype{}, which allowed this loophole to be caught. Fixing the problem is as simple as adding a further instruction in the system prompt to prohibit this behavior. This highlights that the system's context acts as a governance layer for the LLM to correctly fulfill the task given to it. A finely tuned system prompt is thus an important requirement for any system that outsources Cypher generation to an LLM.

In several interactions, Neo4j-related error messages were displayed. These errors were particularly associated with attempts by the LLM to write to the database. The current system architecture prevented permanent modifications to the database during the workshop session, as all Cyphers executed by the LLM were done in read-only mode. This design choice was deliberate, as not to allow an uncontrolled modification of the knowledge graph by the LLM. From a research perspective, this restriction was necessary to guarantee data integrity during experimentation and to ensure reproducibility of results. In a more mature version of \prototype{}, write operations could be enabled through controlled interfaces, thereby expanding practical applicability.

A further technical limitation was identified while tasking the LLM to generate an Excel file. While Open AI does support this file generation and output, the frontend and backend handling within \prototype{} does not currently implement the necessary mechanisms to receive, store, and transmit such artifacts. This means that even if the LLM successfully generated the file server-side, the system architecture lacked the capability to display the results to the user. The limitation is therefore not rooted in the LLM's reasoning capability, but in missing infrastructural components. It highlights that the prototype primarily focuses on query generation and knowledge retrieval rather than comprehensive document management functionality.

Even when no errors were being thrown, \prototype{} still displayed further technical shortcomings. One such moment was identified during the chain of prompts requesting the business support matrix. Even after multiple, refined prompts requesting a matrix, the system would not comply and instead responded with unrequested analysis. This case highlights that an LLM may not inherently understand the operational objective behind a request. In order to fulfill a concrete task, it must therefore be meaningfully guided through structured instructions and clearly defined expectations. While this level of orchestration was manageable during the academic review, it demonstrates that the current system requires deliberate steering to reliably produce task-oriented outputs in a real-world context.

This can be put in contrast to Rovo's behavior, which did not throw any errors during the final review and demonstrated technical stability throughout the process. The system consistently produced coherent outputs without requiring manual intervention or refined steering. This level of robustness indicates a higher degree of technical maturity and integration readiness.

The above results demonstrate a level of technical immaturity within \prototype{}. However, within the scope of this research, a certain degree of technical immaturity was expected and accepted. \prototype{} was conceived and evaluated as a proof-of-concept to validate architectural principles. It is not meant as a production-ready enterprise tool. Qualitative aspects such as stability represent a separate development trajectory from the core research objective of transparent knowledge graph interaction. Within that objective, functional validity was successfully demonstrated, even if technical robustness remains incomplete. These shortcomings would typically be addressed in a subsequent engineering phase and therefore constitute opportunities for future work rather than contradictions of the research objectives.

\section{Make or Buy - Comparing \prototype{} to Rovo}
\label{sub:makeOrBuy}
To do: \cite[pgs. 84-87]{jung2021masterclass} Masterclass EAM Seiten 84-87 einbauen hier.

Although the prototype is not meant as a mature system ready to be applied in real-world scenarios, it serves well as a discussion point for a typical make-or-buy decision when questioning whether licensing a RAG-based LLM tool or building a custom tool is more advantageous for an enterprise. The argument can be framed as buying a black-box system versus developing one's own grey-box system. The decision is therefore both technical and strategic.

A central advantage of a custom tool such as \prototype{} lies in its ability to incorporate expert knowledge via RAG. The knowledge graph containing the entirety of an EAM textbook allowed generated answers to be grounded  in explicit domain knowledge. This grounding mechanism enables a system to expand its knowledge beyond the pretrained general knowledge of an LLM. A system such as Rovo cannot achieve this level of domain specificity without access to a proprietary knowledge graph, as it relies on the information of the pretrained model and the information available in Confluence. While Rovo can summarize and interpret the documented architecture information at hand, it cannot systematically embed structured expert knowledge in the same transparent and controllable manner as \prototype{}. This distinction becomes particularly relevant in specialized domains such as EAM, where correctness and conceptual alignment are critical.

Another decisive advantage of a self-developed system is extensibility. Because the codebase for \prototype{} is fully under internal control, it can be extended to fit any desired use case, given enough time and resources to implement such changes. The risk of extending \prototype{} is also rather low, as the codebase remains manageable in size, modular in structure, and fully under internal control, allowing new features or architectural adjustments to be implemented without dependency on an external provider. Through the MCP server, further systems can be connected in a standardized manner, allowing integration beyond \prototype{}'s scope. In contrast, Rovo is constrained by the functionality offered by Atlassian. Enterprises utilizing Rovo are dependent on the vendor's changes and cannot directly modify or extend the system's internal logic.

This extensibility directly relates to the risk of vendor lock-in. When relying entirely on a third-party system such as Rovo, an enterprise becomes dependent on external decisions regarding pricing models, features, and product strategy. A custom-built solution such as \prototype{} mitigates this dependency by preserving architectural control. The enterprise retains control over its data, interfaces, and system evolution, with the exception of the underlying LLM provider.

However, the final review clearly demonstrated that Rovo currently represents the more mature solution. The system generated responses in a consistent and structured format, often restating the question and summarizing the relevant context before providing an answer. This behavior was positively noted by the experts, as it signaled correct interpretation of the task and ensured clarity. Rovo operated without visible system errors and required less interaction to steer. In contrast, \prototype{} required more deliberate prompting and exposed instabilities that would not be acceptable in a production environment. From a usability and stability perspective, Rovo therefore demonstrated a higher degree of operational maturity.

At the same time, this maturity comes at the cost of transparency. Rovo does not allow inspection of how information is retrieved or how conclusions are derived. In cases of incorrect or misleading outputs, the user has no means to retrace the reasoning path. While Rovo was not immune to hallucinations, its black-box nature limits the ability to detect and analyze them. In contrast, \prototype{} exposes the generated Cypher queries and the underlying raw graph data, as previously discussed, allowing answers to be verified against their source. This transparency shifts responsibility to the user by enabling  traceability, which may be a critical point in enterprise contexts.

The make-or-buy decision therefore becomes a trade-off between maturity versus control. Rovo represents a stable, polished, and immediately deployable solution. \prototype{} represents a flexible, transparent, and extensible architecture that would require further engineering effort to reach comparable stability. Enterprises must therefore decide whether they prioritize convenience and abstraction, or architectural control and transparency.

\section{Practical Applicability in EAM}
\label{sub:practicalApplicability}
\prototype{} has implications for how enterprise architects in general interact with architectural data. Rather than manually using modeling tools, the enterprise architect is able to formulate a task in natural language and receive a structured response grounded within the architectural data. This reduces the friction when information is present in the architecture repository, but difficult to find or extract. During the workshop, experts used the generated outputs as a basis to describe and reflect on the actual state of the landscape. This indicates that the system does not merely retrieve data, but can actively support architectural analysis.

Looking at the work of a junior level enterprise architect, this enables them to deal with data that may otherwise be out of scope for their skill level. With the support of such a system, a junior architect may be able to gain an understanding of the data and develop a course of action for the architecture that would otherwise not be possible as a junior.

The discussion during the development also raised the question of whether a graph database is strictly necessary for such a system. It was argued that in smaller scopes, an LLM may be able to construct contextual relationships in a more implicit, black-box manner without a formalized graph structure. However, once a certain level of scale and complexity is reached, explicit structuring of relationships becomes essential. In such cases, the graph database does not only serve as knowledge base, but as a transparent and inspectable foundation for reasoning. The combination of a knowledge graph and an LLM therefore becomes particularly relevant in enterprise contexts, where traceability, scalability, and avoidance of vendor lock-in are decisive factors.

\section{Learnings From the Implementation}
\label{sub:implementation}
Many learnings can be extracted from the implementation of \prototype{}. This process revealed that the implementation of a RAG-based system should not be overly reliant on LLMs. In early cycles, several core functions for transforming the data were delegated to the LLM. While this accelerated prototyping, it introduced instability and unpredictability in core system components. Replacing LLM-based parsing with deterministic APOC procedures significantly improved reproducibility and trust in the resulting knowledge graph. This shows that LLMs are powerful for reasoning and language generation, but it should be considered to implement critical infrastructure components with deterministic mechanisms.

A further learning from the development phase concerns schema awareness. Pre-querying the database for its schema so that Cyphers can be generated dynamically proved to be a major architectural improvement. Hardcoding the schema into the system prompt was sufficient as long as the dataset remained stable. However, once heterogeneous datasets were introduced, such as when users were able to import their own XML files, this approach no longer scaled well. The shift toward retrieving the current schema via \path{apoc.meta.schema()} and injecting it into the system context allowed the prototype to elevate to a level where it could successfully answer all three categories of questions. This allowed the LLM to operate on an up-to-date representation of the graph structure at runtime. This also demonstrates that text-to-Cypher generation is only robust when the LLM has access to a representation of the knowledge graph. Schema retrieval is therefore a core feature within \prototype{}.

A final key learning concerns the importance of the system prompt as a control mechanism for the LLM. Throughout the development cycles and especially during the final review, it became evident that the system prompt effectively defines the behavioral guard rails of the LLM. The observed Cypher trick was not a failure of the LLM, but a consequence of insufficiently constrained instructions. This demonstrates that the system prompt is not merely contextual guidance, but also a control layer that allows a model to better interpret the task at hand. Therefore, designing, refining, and testing the system prompt must be considered a critical activity.

In summary, the implementation demonstrated that building a transparent, schema-aware, and extensible LLM system requires a deliberate architecture to enable this. LLM capability alone is not sufficient to guarantee a fully functional prototype. Only through deterministic ingestion, dynamic schema awareness, and controlled query generation can a system such as \prototype{} operate in a way that allows users to interact with the knowledge graph via natural language while remaining both inspectable and adaptable to enterprise environments.

\section{Overall Goal Fulfillment}
\label{sub:overallGoalFulfillment}
With the conclusion of the final review and the preceding discussion of the results, a retrospective comparison between the expected outcome from the beginning of the research and the current state of \prototype{}  is appropriate.

Before the first development cycle commenced, it was unclear what the final system was to look like. From vague definitions in the beginning of a chatbot that allows enterprise architects to interact with their data via natural language, it was unclear how this goal would be achieved. With each iterative development step during the conducted Action Research, the end goal became increasingly clear. The result of this is the final version of the prototype.

The architectural objectives of the research can be considered fulfilled. An end-to-end pipeline was successfully implemented, enabling the ingestion of architectural data, its representation within a knowledge graph, the generation of Cypher queries from natural language prompts, and the transparent retrieval of results. The ability to inspect generated Cyphers and validate the underlying data directly in Neo4j demonstrates that the intended grey-box design was achieved.

At the same time, the final review revealed clear limitations in terms of maturity. The system requires structured prompting, is susceptible to hallucinations, and does not yet provide the robustness expected from a production-ready enterprise solution. These shortcomings, however, do not contradict the research contribution. Rather, they highlight that architectural feasibility and operational stability are separate dimensions. The prototype validates the expected concept, even if further engineering effort would be required for a more mature deployment.

Overall, \prototype{} fulfills its primary objective as a proof-of-concept artifact within an Action Research setting. It demonstrates that natural language interaction with a proprietary EAM knowledge graph is technically feasible and allows for transparent and extensible interaction mechanisms. The remaining gaps lie not in conceptual viability, but in the path from prototype to mature enterprise system. Therefore, the core goal of this research, whether a transparent, RAG-based interaction mechanism for EAM data can be realized, is therefore answered positively.


% Conclusion Paragraph
In summary, this chapter has shown that the prototype succeeds architecturally while remaining operationally immature. It highlights both the potential and the structural constraints of integrating LLMs with enterprise knowledge graphs. The findings demonstrate that transparency, schema awareness, and deterministic infrastructure are decisive design principles when developing a RAG-based chatbot system in the realm of EAM. These results form the foundation for the concluding reflections and future research directions discussed in the next chapter.

