\chapter{Conclusion and Future Work}
\label{ch:conclusion}
% Restate problem context
Enterprise Architecture Management requires the analysis and maintenance of large amounts of interconnected architectural information, which is often distributed across heterogeneous documentation formats and tools. Retrieving and editing the relevant information is a challenge of its own for enterprise architects, as this must be done manually and requires a certain level of experience. AI chatbots can support this process by allowing the enterprise architect to interact with the data through natural language. However, such chatbots are typically black-box systems and introduce problems of their own, which may hinder their adoption in real-world settings.

% Research objective and approach
The objective of this research was to investigate whether a transparent RAG–based chatbot can support enterprise architects in accessing proprietary architectural knowledge through natural language interaction. The result is the finished prototype \prototype{}. Using an LLM to generate queries based on the user's natural language input, it is able to retrieve relevant data from a proprietary knowledge graph in order to supplement its responses. \prototype{} was created through Action Research, in which the researcher, along with two practitioners, iteratively developed the system across four cycles. Throughout each cycle, the current problems were diagnosed, action plans were made, and actions were taken before evaluating and specifying the learnings from the cycle. This enabled a collaborative, exploratory approach.

% Main Findings
The results of the conducted Action Research cycles and the final evaluation demonstrate that natural language interaction with a proprietary EAM knowledge graph is technically feasible. \prototype{} demonstrates this through its end-to-end pipeline, allowing a user to upload proprietary enterprise architecture data into the knowledge graph and subsequently querying it. The retrieval process is done by transforming the user's input into a Cypher query using an LLM. This query is then executed within the Neo4j database. The LLM's natural-language response to the user is then augmented with information retrieved from the database. Along with the response, the user is also returned the exact Cypher used, allowing them to manually execute it within Neo4j. This enables the user to inspect the results and ensure factual correctness of the LLM's response.

% Takeaway
One of the central insights of this research is the importance of transparency mechanisms when integrating large language models with enterprise knowledge systems. A central problem when working with LLMs is their hallucinated responses. By showing which Cypher is used to generate the response, \prototype{} is able to mitigate its hallucinations, as a user is able to validate the factual correctness of the generated response. At the same time, the final review showed that the quality of generated answers depends strongly on the prompt context and the specialist terminology used by the user. Without sufficient context, the system does not reliably infer the correct link between the user's prompt and the corresponding part of the knowledge graph. This further highlighted that schema awareness and a well-designed system prompt are necessary in order for the LLM to generate meaningful Cyphers for a heterogeneous knowledge graph. Another central insight is that critical system functions, such as data ingestion and schema handling, should not be left entirely up to the LLM but require deterministic mechanisms in order to achieve robustness and reproducibility. In this way, the research showed that the value of such a system lies not only in natural language interaction with enterprise architecture data, but in doing so in a way that remains inspectable, controllable, and extensible. These insights emerged throughout the development and final review, justifying the applied Action Research method.


%%%%%%%
%%%%%%% Future Work
%%%%%%%
\section{Future Work}
\label{sub:conclusion:futureWork}
Despite the promising results, the developed prototype remains a proof of concept artifact rather than a production-ready enterprise solution. The following discusses possible improvements and future work on \prototype{}.

% Improving prototype robustness
Several technical improvements could further increase \prototype{}'s robustness and usability. For example, the system is very sensitive to its prompt context. Ambiguous or loosely formulated prompts can lead to the generation of Cypher queries that do not target the correct node types within the knowledge graph. Improving the schema-awareness of the prompting strategy or introducing intermediate validation steps could help reduce this dependency on prompt formulation.

Another limitation concerns the use of the database schema within the prompt context. Currently, the schema is supplied before each prompt in order to guide the LLM in generating valid Cypher queries. While this approach improves correctness, it may become problematic as the knowledge graph grows. Prior research has shown that providing extensive schema descriptions can overload the LLM's context window and negatively impact performance. \cite{wan2025prompting, mihindukulasooriya2023text2kgbench} Future work could therefore investigate more efficient methods for supplying schema information.

Furthermore, the data parsing used for transforming Archi XML exports into the knowledge graph may add unnecessary information to the graph. For example, the exported XML files from Archi contain not only the architectural elements themselves but also additional metadata related to visualization and layout. Currently, this information is not filtered out during parsing. Investigating whether this data introduces noise into the knowledge graph and refining the parsing accordingly could further improve retrieval quality.

% Improving deployment for the real world
Beyond improving the technical robustness of the prototype, future work should also address the requirements for deploying \prototype{} in real-world enterprise environments.

A major requirement in such environments is the management of access rights to sensitive information stored in the knowledge graph. This compliance requirement falls into the responsibilities of an enterprise architect. In practice, not every user should be allowed to view all architectural elements or related documentation. Future work should therefore investigate how role-based access control can be integrated into the interaction pipeline, ensuring that generated Cypher queries and returned answers are constrained by the permissions of the respective user.

Another relevant improvement concerns the handling of artifacts generated from system responses. During the final review, it became apparent that downloadable files generated from OpenAI responses, such as Excel files, would enhance the system's practical usefulness. Extending the current frontend and backend so that generated files can be received, stored, and made available for download would therefore be a meaningful step toward a more complete application workflow.

In addition, the current interaction flow could be extended by replacing the largely direct prompt-to-query process with a more explicit pipeline of intermediate steps. Rather than relying on a single generation step to interpret and fulfill the prompt, future work could investigate orchestration-based approaches in which the question is first analyzed, broken down into smaller subtasks, and processed step by step. Such a pipelined interaction strategy may improve robustness and make the system’s internal processing more controllable.

Because \prototype{} already uses MCP as a standardized interface to external systems, future work could also explore connecting it to additional MCP-enabled applications. For example, instead of limiting the system to retrieving and explaining architectural information, \prototype{} could be extended to interact directly with tools such as Archi or Confluence. This would allow the system not only to analyze architectural data but also to support follow-up tasks, such as updating documentation. Such extensions would move the system further from a question-answering prototype toward a broader architecture support system.

% Conclusion
Overall, this research demonstrates that transparent interaction with enterprise architecture knowledge through a RAG-based system is technically feasible. By combining large language models with a structured knowledge graph and exposing the generated queries used for retrieval, \prototype{} provides a grey-box interaction mechanism that reduces the opacity typically associated with AI-based systems. While the prototype represents only an initial step toward practical deployment, the results indicate that a transparent, schema-aware chatbot system can play a meaningful role in supporting enterprise architects as they navigate complex architectural landscapes.


