\chapter{Introduction}
\label{ch:intro}

% Problem context and motiviation - The world has a problem.
In the field of  Enterprise Architecture Management (EAM), artifacts containing architectural information can require large manual effort to understand, extract information from, and update. This is due to the various data sources within an enterprise's architecture. Information is strayed across heterogeneous diagrams and documentation tools. This requires the enterprise architect to manually sift through these sources in order to retrieve the information that they need. Once retrieved, they must attempt to understand what the current state of the architecture is and how its different parts of the architecture relate to one another. Finally, once the data is extracted and understood, it must be updated in order to achieve a certain goal. Extracting, understanding, and updating the architecture can be a daunting task, especially for novice enterprise architects, as this requires understanding the architecture at a detailed level. \cite{jung2021masterclass} This increased effort introduces the risk of inconsistencies, artifacts which are hard to maintain and scale, and hinders data-based decision making.

Through recent advances in artificial intelligence (AI) and Large Language Models (LLMs), it has become possible to process natural language and leverage this to solve tasks. LLMs are capable of interpreting user inputs formulated in natural language in order to generate structured responses. This is achieved by training the LLM on large amounts of data. \cite{singhal2023large, hadi2023large} In combination with retrieval mechanisms, such as Retrieval-Augmented Generation (RAG), it is possible to supplement the responses of an LLM with the information of a specific knowledge base. This augmentation enables the system to answer domain-specific questions by retrieving relevant information and using it as context for answer generation. \cite{zhao2026retrieval} Such systems are often implemented as chat-based interfaces, allowing users to interact with the stored knowledge through natural conversation. This opens up the possibility of supporting enterprise architects in navigating and understanding complex architectural landscapes, reducing manual effort and increasing the speed at which the data can be understood.

% Research gap - Existing solutions are insufficient.
Academic research has extensively explored the integration of LLMs with structured knowledge sources such as knowledge graphs in order to reduce hallucinations and improve answer quality. Furthermore, research highlights the importance of grounding LLM outputs in domain-specific data when applying them to critical fields. \cite{lewis2020retrieval, sharma2025reducing} However, there is limited work that demonstrates how such approaches can be applied specifically within the domain of Enterprise Architecture Management using proprietary architectural data and evaluated in a real-world setting. In particular, existing research rarely addresses how transparency can be maintained when an LLM dynamically interacts with enterprise architecture data. Also, there is little documented evidence of how such a system can be iteratively validated together with domain experts.

Commercial solutions also exist. However, these solutions generally come with the trade off of being operated by third parties, thus leaving control up to these providers. Because the details of such a system are rarely open source, the system becomes a black-box, disabling an enterprise architect from understanding how the tool reasoned, which data was retrieved, and how the final answer was generated. This lack of transparency can be particularly problematic when dealing with proprietary and business critical architectural information. Without transparency, the line between retrieved facts, inferred relationships, and hallucinated content by the LLM becomes increasingly difficult to distinguish.

What is currently missing is a transparent, RAG-based chatbot, backed by a proprietary EAM knowledge graph, that explicitly exposes its retrieval and query mechanisms and is developed and evaluated in close collaboration with EAM practitioners.

% Research goal (in place of the question) - This thesis addresses that gap.
Thus, the goal of this thesis is to develop and evaluate a prototypical RAG-based chatbot that supports tasks in EAM. The purpose of the chatbot is to support junior level enterprise architects in extracting and understanding architectural data better, allowing them to update the architecture without the otherwise required detailed understanding of the landscape. The chatbot is connected to a knowledge graph which has textbook information on EAM stored within, enabling the LLM to act as an EAM-expert. The knowledge graph also contains the proprietary enterprise architecture data. A user is able to interact with the chatbot and ask it questions regarding the stored data. A key feature within the chatbot is the aspect of transparency, which is ensured by allowing the user to examine the section of the knowledge graph used by the LLM to augment its answer.

% Methodological overview - It uses this method.
Action Research was applied as the methodological approach, as this enables an explorative development approach while maintaining academic integrity. This allowed the researcher and two expert practitioners to iteratively develop the prototype and guide the development in the right direction by assessing the system’s functionality, usability, answer quality, and alignment with practical EAM requirements in each cycle. After the final iteration, a final review was conducted in order to evaluate the state of the system and whether or not the goal of the research was achieved with the prototype.

% Contribution of my thesis - It contributes the following insights.
The research at hands contributes to the research domain in various ways. On a technical level, it demonstrates how a RAG-based chatbot can be implemented using a proprietary EAM knowledge graph and how LLM-generated queries can be used to extract relevant information from it. Furthermore, it illustrates how retrieval and query mechanisms can be exposed in order to increase transparency and reduce the black-box characteristics commonly associated with LLM-based systems. From an EAM point of view, the research highlights the importance of transparency and involvement of practitioners when applying an LLM to an enterprise's architectural data in order to maintain integrity within the architecture. The research also shows how such a system can be iteratively developed and evaluated in collaboration with domain experts.

% Thesis Structure
This thesis is structured as follows. Chapter \ref{ch:background} defines the key concepts used throughout the research. Chapter \ref{ch:SOTA} details how research was conducted, gives an overview of current research as well as implementations within the domain. Chapter \ref{ch:actionResearch} describes what the Action Research method is, while the subsequent chapters \ref{ch:cycle1}, \ref{ch:cycle2}, \ref{ch:cycle3}, and \ref{ch:cycle4} describe the details of each iteration of the development. Chapter \ref{ch:finalReview} explains how the final review was conducted and what was observed. Building on the results of the development cycles, chapter \ref{ch:implementation} explains the final prototype's architecture and technical details. Chapter \ref{ch:Discussion} then discusses the results of the Action Research and what the results imply. Finally, chapter \ref{ch:conclusion} presents the conclusion and possibilities for future work.

This thesis assumes an understanding of general terms of the reader in the realm of application development and enterprise architecture management. Understanding at an intermediate level is sufficient, as all key concepts are explained in detail in the next chapter.